---
title: "Yield monitor wrangling - 2017"
author: "All of us"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Learning objectives  
Today's learning objectives are:  

- Talk about the role of yield monitor data in ag  
- Import yield data  
- Explore with tables, plots, and maps (EDA)  
- Learn and implement data wrangling steps for geospatial data  

# Setup  
## Loading packages  
```{r setup }
#| message: false

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)

```

## Importing data  
```{r yield17 }
yield17 

yield17 

```

# EDA  
```{r summary}
summary(yield17) 
```

What columns do you think are relevant?  

```{r yield distribution - density plot}
ggplot(yield17, aes(x=Dry_Yield))

```

```{r speed distribution - density plot}
ggplot(yield17, aes(x=Speed))+
  geom_density()+
  geom_rug()

```

```{r yield map}


```

```{r yield restricted map}


```

```{r yield x speed}
ggplot(yield17, aes(x=Speed, 
                    y=Dry_Yield))

```
Do we need to clean this data?  
From a yield map quality perspective, what steps would you do?  

- Clean based on speed  
- Clean based on yield  

```{r data needs}
yield17
```

From a data wrangling perspective, what steps would you do?  

- standardize column naming style (for ease of use)  
- select only needed columns:  

  - Speed  
  - Dry_Yield  
  - POINT_X  
  - POINT_Y  

- rename columns to include measurement units  
- create columns in metric system  
- transform it into spatial object (as of now, it is only a data frame)  
- transform from geographic (unprojected) to projected  

# Wrangling  
```{r yield17_w}
yield17_w <- yield17 %>% 
  # Cleaning names
  clean_names() 
  # Selecting desired variables and renaming some of them
  
  # Transforming from english to metric units
  #mutate(yield_kgha=yield_lbac*0.453592/0.404686,
  #       speed_kmh=speed_mph*1.60934)

  # Transforming into sf object

  # Setting native crs (geographic)

  # Transforming from geodraphic to prjected CRS

yield17_w

```

# Maps  
```{r static}
yield17_w %>%
  filter(yield_lbac < 3000) %>%
  ggplot()+
  geom_sf(aes(color = yield_lbac), size = .1)+
  scale_color_viridis_b()

```

```{r interactive}
yield17_w %>%
  filter(yield_lbac < 3000) 

```

# Quiz  
Where in the field do you see low-yielding areas? Of these different low-yielding areas, which ones do you think can be related to yield monitoring errors? Why?  

Tip: Toggle the yield layer on and off above, it may help you to answer.  

You have 5 min. Send me your answer by email (lmbastos@uga.edu). This will also be your attendance.  



# Exporting  
```{r}


```


