---
title: "Yield monitor editor - 2017"
author: "All of us"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: sketchy
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Learning objectives  
Today’s learning objectives are:  

- Learn and implement yield monitor data cleaning steps  
- Compare numerically and graphically raw and cleaned yield monitor data

# Setup  
## Loading packages  
```{r  }
#| message: false

# installing 
#install.packages("leafsync")
#install.packages("patchwork")

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)
library(leafsync)
library(patchwork)

```

## Importing data  
```{r }
yield17_w 

yield17_w 
```

# EDA  
## Summary  
```{r }
summary(yield17_w) 
```

## Raw map
```{r}
yieldmap_raw <- ggplot(yield17_w)

yieldmap_raw
```

## Raw map interactive  
```{r}
#| warning: false

yieldmap_raw_int <- yield17_w %>%
  filter(yield_lbac < 3000) %>%
  mapview(zcol = "yield_lbac",
        col.regions = viridis(3),
        alpha.regions = 1,
        cex = .1
        ...,
        ...
)

yieldmap_raw_int
```

Today we are applying 5 filters:  

- Min and max yield  
- Min and max speed  
- Position  

# Yield editor - Min/max yield  
## Raw yield distribution  

```{r}
ggplot(data = yield17_w, 
       aes(x = yield_lbac))+
  geom_density()+
  geom_rug()
```

## Raw yield summary  
```{r}

```

## Raw yield quantiles  
```{r}
lowq_yield 
lowq_yield

upq_yield 
upq_yield

iqr 
iqr

miny 
miny

maxy 
maxy
```

## Min/max raw yield quantiles plot  
```{r}
ggplot()+
  geom_density(data = yield17_w, 
               aes(x = yield_lbac))

```

## Filtering min/max raw yield  
```{r}
yield17_w_yf 

yield17_w_yf
```

Before filtering we had a total of 31,724 data rows.
How many rows now after filtering?

## Min/max yield cleaned summary  
```{r}

```
All right, all done filtering based on yield min and max.
Now let’s do the same for speed.

# Yield editor - Min/max speed  
## Raw speed distribution  
```{r}
ggplot(data = yield17_w, 
       aes(x = speed_mph))+
  geom_density()+
  geom_rug()
```

## Raw speed summary  
```{r}
summary(yield17_w$speed_mph) 
```

## Raw speed quantiles  
```{r}
lowq_speed 
lowq_speed

upq_speed 
upq_speed

iqr <- upq_speed - lowq_speed
iqr

minv <- lowq_speed - (0.18*iqr)
minv

maxv <- upq_speed + (0.18*iqr)
maxv
```

## Min/max raw speed quantiles plot  
```{r}
ggplot()+
  geom_density(data = yield17_w, 
               aes(x = speed_mph))+
  geom_vline(xintercept = lowq_speed, 
             color = "red", 
             linetype = 2)+
  geom_vline(xintercept = upq_speed, 
             color = "blue", 
             linetype = 2)+
  geom_vline(xintercept = minv, 
             color = "red", 
             linetype =1)+
  geom_vline(xintercept = maxv, 
             color = "blue", 
             linetype = 1)+
  annotate(geom = "rect",
           ymin = 0, 
           ymax = Inf,
           xmin = minv, 
           xmax = maxv,
           alpha = .6, 
           fill = "green")

```

## Filtering min/max raw speed  
```{r}
yield17_w_sf 

yield17_w_sf

```

## Min/max speed cleaned summary  
```{r}

```

# Yield editor - Position  
Original position filter was based on quantiles too.
Why filter based on boundary?  

## Field boundary  
```{r}
boundary 

boundary
```

```{r}
boundary_w 
  
boundary_w
```

## Field boundary plot  
```{r}
yieldmap_raw
```

Let’s create a negative buffer of 20 m.  

## Buffer  
```{r}
buffer 

buffer
```

```{r}
yieldmap_raw + 
  geom_sf(data=boundary_w, 
          fill = NA)
```


## Filtering position  
```{r}
yield17_w_pf 

yield17_w_pf
```

## Position yield cleaned summary  
```{r}
summary(yield17_w_pf$yield_lbac)
```

# Raw vs Cleaned data  
## Summary statistics  
### Raw yield summary  
```{r}
summary_raw 

summary_raw
```

### Clean yield summary  
```{r}
summary_filtered <- ... %>%
  summarise(n = length(yield_lbac),
            min = min(yield_lbac),
            mean = mean(yield_lbac),
            max = max(yield_lbac),
            sd = sd(yield_lbac),
            cv = (sd/mean)*100
            ) %>%
  mutate(data = "clean") %>%
  dplyr::select(data, everything())

summary_filtered
```

### Combined yield summary  
```{r}
summary_raw %>%
  bind_rows(summary_filtered)
```

## Density plot  
```{r}
ggplot()+
  geom_density(data = yield17_w, 
               aes(x = yield_lbac), 
               fill = "blue", 
               alpha = .5)
```

# Maps  
## Raw yield map  
```{r}
yieldmap_raw
```

## Clean yield map  
```{r}
yieldmap_clean <- ggplot(yield17_w_pf)+
  geom_sf(aes(color = yield_lbac), 
          size = .1)+
  scale_color_viridis_b()

yieldmap_clean
```

## Clean yield map interactive  
```{r}
yieldmap_clean_int <- yield17_w_pf %>%
  mapview(zcol = "yield_lbac",
        col.regions = viridis(3),
        alpha.regions = 1,
        cex = .1,
        lwd = 0,
        map.types = "Esri.WorldImagery"
)

yieldmap_clean_int
```

## Side-by-side raw vs. cleaned yield map  
```{r}
yieldmap_raw + yieldmap_clean
```

## Side-by-side raw vs. cleaned yield map interactive  
```{r}

```

# Exporting cleaned version  
```{r}


```

# Summary  
Today we:  

- Learned about and implemented 5 yield filters (yield, speed, position)  
- Compared raw and cleaned yield monitor data

