---
title: "Soil ECa"
author: "All of us"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Learning Objectives  

Today, our objectives are to:  

- Import soil ECa data  
- Interpolate soil ECa point data for two depths   
- Export soil ECa layers to file  

# Setup  
```{r setup }
# Packages
library(dplyr)
library(tidyr)
library(janitor)
library(readr)
library(sf) #vector manipulation
library(ggplot2)
library(viridis)
library(ggthemes)
library(gstat) #IDW
library(stars) #raster manipulation
library(terra) #topographical variables
library(patchwork) #combining multiple plots  
library(tibble) #rownames to column

```

```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "right",
        panel.background = element_rect(fill="gray80",
                                        color=NA),
        plot.background = element_rect(fill="gray80",
                                       color=NA),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))

```

```{r reading elevation data}
eca 

eca
```
# Wrangling  
```{r eca_w}
eca_w 

eca_w 
```

# EDA  
```{r eca30 plot}
eca_w 
```


```{r eca90 plot}
eca_w 

```

# Grid  
Since we are interpolating, we need to create a grid (same as for the yield and terrain exercises) based on the field boundary.    

```{r boundary_w}
boundary_w 

boundary_w
```

```{r grid_r}
# grid in vector format
grid_r 

grid_r

ggplot()+
  geom_stars(data = grid_r)+
  mytheme_map

```

# Interpolation
## eca30  
```{r eca30_idw_mod}
eca30_idw_mod 

eca30_idw_mod
```

```{r eca30_idw_pred}
eca30_idw_pred  

eca30_idw_pred
```

```{r eca30_map }
eca30_map 

eca30_map
```

## eca90  
```{r eca90_idw_mod}
eca90_idw_mod 

eca90_idw_mod
```

```{r eca90_idw_pred}
eca90_idw_pred  

eca90_idw_pred
```

```{r eca90_map }
eca90_map <- ggplot()+
  geom_stars(data = eca90_idw_pred)+
  scale_fill_viridis_c()+
  geom_sf(data=boundary_w, fill=NA)+
  labs(fill="ECa\n0-90 cm\n(dS/m)")+
  mytheme_map

eca90_map
```

## Plotting all together
```{r plotting all together}
eca30_map + eca90_map

ggsave("../output/soileca.png", 
       width = 6, 
       height = 3)
```

# Grid extraction  
```{r eca30_v}
eca30_v 
```


```{r eca90_v}
eca90_v <- eca90_idw_pred %>%
  st_as_sf(as_points = F, 
           merge = F) %>%
  dplyr::select(eca90_dsm=var1.pred) 

eca90_v
```

# Merging eca layers  
```{r eca_v }
eca_v 

eca_v
```

# Exporting  
Exporting to file to use later for management zone creation.  
```{r exporting}


```

# Summary  
Today we have:  
- Imported point vector soil ECa data  
- Interpolated soil ECa at two different depths to same grid as used for yield (keeping them all on same spatial extent and resolution)  
- Transformed them into vector, joined both depths in one object, and exported to file  

# What's next?  
Next we will be ready to create and validate management zones!

For that, we will first load all layers created until now together.

Then, we will check their relationships, run a tree algorithm to explore variable importance (more on that later), and finally use a clustering algorithm to create zones. 